Landau quantization of nearly degenerate bands

### 2018-10-01 paper.tex
Updates from ShareLaTeX

The quantization rule for Landau levels reflects the single-valuedness of the semiclassical (WKB) wavefunction over $\frako_0$,[Berry, Mount] i.e.,  the phase accumulated by an electron (over one cyclotron period $T_c$) is an integer multiple of $2\pi$: $(l^2S(E){+}\gamma)/2\pi{\in}\Z$. The leading-order phase   involves the classical action $S{:}{=}-\int_{0}^{T_c} k_x \dot{k}_y dt$  corresponding to $(K_x,K_y)$;[mount] $S$ has a dual interpretation as the signed area (in $\bk$-space) enclosed by $\frako_0$, as illustrated in Fig. \ref{fig:orbits}(b). $\gamma$ is a subleading Maslov correction [caustic] that generally equals $\pi$ times the rotation number $r$ of $\frako_0$ (i.e., the number of rotations made by the tangent vector over one cycle);[100page] for orbits that are deformable to a circle,  $r{=}{\pm} 1$. In the zeroth-order  solution, all Landau levels are spin-degenerate with a level spacing 
\e{ \var_c:=\f{2\pi}{l^2|\partial S/\partial E|}:=\f{\hbar^2}{ml^2}:=\f{h}{T_c}. \label{def:cyclotron}} 
Here, $\var_c$ and $T_c$ are respectively the cyclotron energy and period of the zeroth-order orbit, and $m$ is the effective mass. Henceforth, `degeneracy of Landau levels' should be understood as a spin degeneracy.

@@

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta \hat{H}$ to the degenerate band subspace of $\hat{H}_0$. In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively), $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,\delta \var_-$, corresponding to outer and inner orbit respectively; $\delta \var_+<\delta \var_-$) are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of  \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \label{spinsplitwkb}}
%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics[nenciu] along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for quasidegenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \label{phaseindependentorbit}} 
In the second equality, we have applied \q{spinsplitwkb} and assumed that 
 $\delta \var$ is traceless (this assumption is  valid for splittings of a spin degeneracy\footnote{The Pauli spin-orbit coupling term ${\propto} \bs{\cdot} (\bE {\times} \bp)$ (that is derived from the non-relativistic limit of the Dirac equation) is traceless in any spin-symmetric (pseudo-)spinor basis.  Viewing the spin-orbit coupling as a small parameter in degenerate perturbation theory, the lowest-order splitting of any spin-degenerate energy level is always symmetric about the zeroth-order energy. This follows because the zeroth-order, spin-symmetric wavefunctions (associated to a spin-degenerate level) are tensor products of spin and spatial wavefunctions: $\ket{\Psi_{\pm}}{:}{=}\ket{{\pm}\hat{n}}{\otimes} \ket{\psi}$ with $\hat{n}$ a unit vector and  $\bs{\cdot} \hat{n} \ket{{\pm}\hat{n}}{=}{\pm}\ket{\pm \hat{n}}$. Hence, $\braopket{\Psi_{\pm}}{\bs{\cdot} (\bE {\times} \bp)}{\Psi_{\pm}}{=}{\pm}\hat{n}\cdot\braopket{\psi}{\bE {\times} \bp}{\psi}$. \textbf{Please generalize above proof to the case where H0 already includes centrosymmetric SOC. }}).  \q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule[onsager,lifshitz] for  two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth[rothII].\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S )$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits.[100page] } The perturbative inclusion of $B\calm_{\sma{+-}}$ results in  a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in  \s{sec:qo}.    \\

@@

Due to the continuous rotational symmetry of $H_R$, $\nabk \theta$ has magnitude $1/k$ and lies tangential to $\frako_0$; consequently, $\calh$ is constant along $\frako_0$, and the propagator simplifies to $\cala{=}e^{{-}i\calh T_c/\hbar}$ without time-ordering, and the phase corrections to the quantization rule are simply the eigenvalues of $\calh$ multiplied by $T_c$:
\begin{equation}
\lambda_{\pm}=\pm\f1{2}\sqrt{({l^2\delta S})^2+\pi^2(g_{\sma{\text{eff}}}-2)^2}+\pi. \label{eq:Rashba-lambda}
\end{equation}
This is a function of the differential action $\delta S$ [cf.\ \q{deltaSvsS}] and  the effective  $g$-factor ($g_{\sma{\text{eff}}}{:}{=}g_sm/m_0$), which is the free-electron $g$-factor multiplied with the ratio of  effective ($m$) and free-electron ($m_0$) masses. The standalone $\pi$ term is the single-band Berry phase associated to the winding of $\theta_{\bk}$ by ${\pm}2\pi$. The two quantities under the square root originate from the spin-orbit energy splitting $\delta \var$ and the inter-band matrix elements of $B\calm$ (in the basis of zero-field bands).

In the weak-hybridization regime: $l^2\delta S  \gg \pi|g_{\text{eff}}{-}2|$, we may expand \q{eq:Rashba-lambda} as a Laurent series in $B$:
\e{\lambda_{\pm}= \pm \f{l^2\delta S}{2} +\pi \pm \f{\pi^2}{4}\f{(g_{\text{eff}}-2)^2}{l^2\delta S} + O(B^2). \label{lambdasmallB}}
\q{eq:rule}, combined with the first two terms on the right-hand-side of \q{lambdasmallB}, may alternatively be derived from 
the standard Onsager-Lifshitz rules for two independent orbits, with a common $\pi$ Berry-phase correction for each orbit:\footnote{The intra-band Zeeman and orbital-moment couplings vanish due spacetime-inversion symmetry.[topoferm]} $l^2(S{\pm}\delta S)/2\pi {\in}\Z$. Higher-order terms in \q{lambdasmallB} reflect the field-induced hybridization of the spin-orbit-split orbits.

In the strong-hybridization regime: $l^2\delta S {\ll}\pi(g_{\text{eff}}{-}2)$, $\lambda_{\pm}{\approx}{\pm}\pi g_{\text{eff}} /2$ mod $2\pi$, which corresponds to the Zeeman splitting of Landau levels by $g_{s}\mu_BB$, in the \textit{absence} of the spin-orbit interaction.

@@

Our quantization rule [\q{eq:rule} with \q{eq:Rashba-lambda}] may be compared with the  exact Landau-level spectrum of the Peierls-substituted Hamiltonian $H_R(\bK)$. The exact levels are given implicitly by
\e{
&n_{\pm}=\frac{l^2S}{2\pi}+\frac{l^2\delta S}{16\pi}\f{\delta S}{S}\lin 
&\pm\f1{2}\sqrt{\bigg(\frac{l^2\delta S}{8\pi}\f{\delta S}{S}\bigg)^2+(l^2\delta S)^2+\pi^2(g_{\sma{\text{eff}}}-2)^2},\label{eq:Rashba-exact}}
with the functional forms of $\delta S(E)$ and $S(E)$ provided in \q{deltaSvsS};  $n_{+}{\ge} 1$ and $n_-{\ge} 0$ here are integer-valued labels for two distinct sets of levels.  In the quasidegenerate regime ($E{\gg}m\alpha^2$), terms involving $\delta S/S$ may be neglected and \q{eq:Rashba-exact} reduces to our quantization rule with $\lambda_{\pm}$ given by \q{eq:Rashba-lambda}.

@@

We begin again with the Rashba model of a rotationally-symmetric Dirac cone. To convert this Dirac point to its asymmetric variant, 
let us introduce a Zeeman coupling to an in-plane magnetic field $B_\parallel$ (parallel to the $+y$ direction):
\e{ 
H_{RZ}=\f{{\hbar^2}k^{2}}{2m}+\delta \var; \;\;\;\delta \var=\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\zpar\sigma_{y}, \label{eq:RZ-Hamiltonian}}
with $\zpar{:}{=}{g_s}\mu_B B_{\parallel}/2$.
While this field breaks both time-reversal ($T$) and two-fold rotational ($C_2$) symmetries, the composition $C_2T$ is preserved as $\sigma_x H^*_{RZ}(\bk)\sigma_x{=}H_{RZ}(\bk),$ which constrains spin to lie in-plane at each $\bk$. In this symmetry class, Dirac nodes are irremovable but movable   in two-dimensional $\bk$-space. The in-plane Zeeman field has the dual effect of  moving the Dirac node  to the position:
\begin{eqnarray}
\bar{k}_x & = & -\frac{\zpar}{\alpha},\;\bar{k}_y=0,\;\epsilon_{0}=\frac{{\hbar^2}\zpar^{2}}{2m\alpha^2},\label{whereisdiracpoint}
\end{eqnarray}
and tilting the Dirac cone. To observe this tilt, consider the linearized Hamiltonian in the vicinity of this node: 
\e{
H & =\alpha\left[\left(\sigma_y-\frac{{\hbar^2}\zpar}{m\alpha^2}\right)\delta k_{x}-\delta k_{y}\sigma_{x}\right]+\epsilon_{0}.}
When $|\hbar^2\zpar|>|m\alpha^2|$, the Dirac cone tilts over in the $x$ direction, leading to a Lifshitz transition  of the constant-energy band contours -- from circular to hyperbolic (in the linearized approximation), as illustrated in Fig. \ref{fig:RZ}(c). An over-tilted Dirac cone shall be referred to as type-II. Our case study of a II-Dirac point  connects  two electron-like  pockets [cf.\ \fig{fig:RZ}(a,c)], and  differs from previous case studies [OBrien, AALG] that connect an electron- and hole-like pocket.

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Upper panel: a sketch of Landau level dispersion near II-Dirac point of Rashba model with in plane Zeeman field in the weak hybridization limit. The dashed line representing first order solution [cf. Eq. (\ref{RZfirstorderE})] oscillates around the solid line [zeroth order solution, cf. Eq. ({trefoillerule})]. Due to the smallness of $\delta S$, the oscillation should not be observable in the scale of several Landau levels. Therefore, We exaggerate the oscillation frequency in the sketch. Lower panel: oscillation of $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})] for even integer $n$ with respect to magnetic field calculated by Eq. (\ref{RZfirstorderE}) (black dashed line) and numerical diagonalization (red solid line). $\delta E$ is extracted at 0.45eV and the II-Dirac point is located at 0.5eV. Parameters for lower panel of (b): $m$: 0.076$m_0$, $\alpha$: 1.5($10^{5}$cm/s$\cdot\hbar$), $Z_\parallel$: 1meV. Out of plane Zeeman coupling is ignored here. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: illustration of $S_{\sma{\square}}$ by the purple rectangle. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\cite{RZfignote}, Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$). Parameters for (d): $m=0.076m_0$, $l=100$\AA. \label{fig:RZ}}.
\end{figure}

@@

Away from the II-Dirac point, an electron undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of Landau-Zener tunneling/transmission and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_+}}{e^{i\tilde{\lambda}_-}}. \label{propinplanezeeman}}
The adiabatic phase $\tilde{\lambda}_{\pm}$ is the sum of a dynamical phase ${\pm} l^2 \delta S/2$ and  the single-band Berry phase $\phi^B_\pm$, for the outer ($+$) and inner ($-$) orbit respectively. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and decreasing (resp.\ increasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.

>>>>

The quantization rule for Landau levels reflects the single-valuedness of the semiclassical (WKB) wavefunction over $\frako_0$,[Berry, Mount] i.e.,  the phase accumulated by an electron (over one cyclotron period $T_c$) is an integer multiple of $2\pi$: $(l^2S(E){+}\gamma)/2\pi{\in}\Z$. The leading-order phase   is the classical action $l^2S{:}{=}{-}l^2\int_{0}^{T_c} k_x \dot{k}_y dt$  corresponding to $(K_x,K_y)$;[mount] $S$ has a dual interpretation as the signed area (in $\bk$-space) enclosed by $\frako_0$, as illustrated in Fig. \ref{fig:orbits}(b). $\gamma$ is a subleading Maslov correction [caustic] that generally equals $\pi$ times the rotation number $r$ of $\frako_0$ (i.e., the number of rotations made by the tangent vector over one cycle);[100page] for orbits that are deformable to a circle,  $r{=}{\pm} 1$. In the zeroth-order  solution, all Landau levels are spin-degenerate with a level spacing 
\e{ \var_c:=\f{2\pi}{l^2|\partial S/\partial E|}:=\f{\hbar^2}{ml^2}:=\f{h}{T_c}. \label{def:cyclotron}} 
Here, $\var_c$ and $T_c$ are respectively the cyclotron energy and period of the zeroth-order orbit, and $m$ is the effective mass. Henceforth, `degeneracy of Landau levels' should be understood as a spin degeneracy.

@@

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta \hat{H}$ to the degenerate band subspace of $\hat{H}_0$. $\delta \var$ is assumed to be traceless; this property is guaranteed for $\delta \hat{H}$ a spin-orbit coupling that breaks the spin SU(2) symmetry of $\hat{H}_0$;\footnote{The Pauli spin-orbit coupling term ${\propto} \bs{\cdot} (\bE {\times} \bp)$ is traceless in any spin-symmetric (pseudo-)spinor basis.  Viewing the spin-orbit coupling as a small parameter in degenerate perturbation theory, the lowest-order splitting of any spin-degenerate energy level is always symmetric about the zeroth-order energy. This follows because the zeroth-order, spin-symmetric wavefunctions (associated to a spin-degenerate level) are tensor products of spin and spatial wavefunctions: $\ket{\Psi_{\pm}}{:}{=}\ket{{\pm}\hat{n}}{\otimes} \ket{\psi}$ with $\hat{n}$ a unit vector and  $\bs{\cdot} \hat{n} \ket{{\pm}\hat{n}}{=}{\pm}\ket{\pm \hat{n}}$. Hence, $\braopket{\Psi_{\pm}}{\bs{\cdot} (\bE {\times} \bp)}{\Psi_{\pm}}{=}{\pm}\hat{n}\cdot\braopket{\psi}{\bE {\times} \bp}{\psi}$.} more generally, one may always absorb the traceful part (of  the projection of $\delta \hat{H}$) into the definition of $\var(\bk)$.  In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively), $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,~\delta \var_-{=}{-}\delta \var_+$)  are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \label{spinsplitwkb}}

%is not traceless, the traceful part can always be absorbed into the redefinition of $S$ within the accuracy of first order perturbation theory.

%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics[nenciu] along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for quasidegenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \label{phaseindependentorbit}} 
\q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule[onsager,lifshitz] for  two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth[rothII].\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S )$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits.[100page] } The perturbative inclusion of $B\calm_{\sma{+-}}$ results in  a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in  \s{sec:qo}.    \\

@@

Due to the continuous rotational symmetry of $H_R$, $\nabk \theta$ has magnitude $1/k$ and lies tangential to $\frako_0$; consequently, $\calh$ is constant along $\frako_0$, and the propagator simplifies to $\cala{=}e^{{-}i\calh T_c/\hbar}$ without time-ordering, and the phase corrections to the quantization rule are simply the eigenvalues of $\calh$ multiplied by $T_c$:
\begin{equation}
\lambda_{\pm}=\pm\f1{2}\sqrt{({l^2\delta S})^2+\pi^2(g_sm/m_0-2)^2}+\pi. \label{eq:Rashba-lambda}
\end{equation}
This is a function of the differential area $\delta S$ [cf.\ \q{deltaSvsS}] and the effective mass $m$. The standalone $\pi$ term is the single-band Berry phase associated to the winding of $\theta_{\bk}$ by ${\pm}2\pi$. The two quantities under the square root originate from the spin-orbit energy splitting $\delta \var$ and the inter-band matrix elements of $B\calm$ (in the basis of zero-field bands).

In the weak-hybridization regime: $l^2\delta S  \gg \pi|g_s m/m_0{-}2|$, we may expand \q{eq:Rashba-lambda} as a Laurent series in $B$:
\e{\lambda_{\pm}= \pm \f{l^2\delta S}{2} +\pi \pm \f{\pi^2}{4}\f{(g_sm/m_0-2)^2}{l^2\delta S} + O(B^2). \label{lambdasmallB}}
\q{eq:rule}, combined with the first two terms on the right-hand-side of \q{lambdasmallB}, may alternatively be derived from 
the standard Onsager-Lifshitz rules for two independent orbits, with a common $\pi$ Berry-phase correction for each orbit:\footnote{The intra-band Zeeman and orbital-moment couplings vanish due spacetime-inversion symmetry.[topoferm]} $l^2(S{\pm}\delta S)/2\pi {\in}\Z$. Higher-order terms in \q{lambdasmallB} reflect the field-induced hybridization of the spin-orbit-split orbits.

In the strong hybridization regime: $l^2\delta S {\ll}\pi(g_sm/m_0{-}2)$, $\lambda_{\pm}{\approx}{\pm}\pi g_sm/2m_0$ mod $2\pi$, which corresponds to the Zeeman splitting of Landau levels by $g_{s}\mu_BB$, in the \textit{absence} of the spin-orbit interaction.

@@

Our quantization rule [\q{eq:rule} with \q{eq:Rashba-lambda}] may be compared with the  exact Landau-level spectrum of the Peierls-substituted Hamiltonian $H_R(\bK)$. The exact levels are given implicitly by
\e{
&n_{\pm}=\frac{l^2S}{2\pi}+\frac{l^2\delta S}{16\pi}\f{\delta S}{S}\lin 
&\pm\f1{2}\sqrt{\bigg(\frac{l^2\delta S}{8\pi}\f{\delta S}{S}\bigg)^2+(l^2\delta S)^2+\pi^2(\f{g_sm}{m_0}-2)^2},\label{eq:Rashba-exact}}
with the functional forms of $\delta S(E)$ and $S(E)$ provided in \q{deltaSvsS};  $n_{+}{\ge} 1$ and $n_-{\ge} 0$ here are integer-valued labels for two distinct sets of levels.  In the quasidegenerate regime ($E{\gg}m\alpha^2$), terms involving $\delta S/S$ may be neglected and \q{eq:Rashba-exact} reduces to our quantization rule with $\lambda_{\pm}$ given by \q{eq:Rashba-lambda}.

@@

%\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\zpar\sigma_{y},

We begin again with the Rashba model [cf.\ \q{eq:Rashba-Hamiltonian}] of a rotationally-symmetric Dirac cone. To convert this Dirac point to its asymmetric variant, 
let us introduce a Zeeman coupling to an in-plane magnetic field $B_\parallel$ (parallel to the $+y$ direction):
\e{ 
H_{RZ}=\f{{\hbar^2}k^{2}}{2m}+\delta \var; \;\;\;\delta \var=d_x\sigma_x+d_y\sigma_y,\lin
(d_x,d_y):=(-\alpha k_y,\alpha k_x+\zpar),\;\;\;\zpar:= g_s \mu_B B_{\parallel}/2.
\label{eq:RZ-Hamiltonian}}
While this field breaks both time-reversal ($T$) and two-fold rotational ($C_2$) symmetries, the composition $C_2T$ is preserved as $\sigma_x H^*_{RZ}(\bk)\sigma_x{=}H_{RZ}(\bk),$ which constrains spin to lie in-plane at each $\bk$. In this symmetry class, Dirac nodes are irremovable but movable   in two-dimensional $\bk$-space. The in-plane Zeeman field has the dual effect of  moving the Dirac node  to the position:
\begin{eqnarray}
\bar{k}_x & = & -\frac{\zpar}{\alpha},\;\bar{k}_y=0,\;\epsilon_{0}=\frac{{\hbar^2}\zpar^{2}}{2m\alpha^2},\label{whereisdiracpoint}
\end{eqnarray}
and tilting the Dirac cone. To observe this tilt, consider the linearized Hamiltonian in the vicinity of this node: 
\e{
H & =\alpha\left[\left(\sigma_y-\frac{{\hbar^2}\zpar}{m\alpha^2}\right)\delta k_{x}-\delta k_{y}\sigma_{x}\right]+\epsilon_{0}.}
When $|\hbar^2\zpar|>|m\alpha^2|$, the Dirac cone tilts over in the $x$ direction, leading to a Lifshitz transition  of the constant-energy band contours -- from circular to hyperbolic (in the linearized approximation), as illustrated in Fig. \ref{fig:RZ}(c). An over-tilted Dirac cone shall be referred to as type-II. Our case study of a II-Dirac point  connects  two electron-like  pockets [cf.\ \fig{fig:RZ}(a,c)], and  differs from previous case studies [OBrien, AALG] that connect an electron- and hole-like pocket.

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the degeneracy-lifting Hamiltonian $\delta\var$ [cf.\ \q{eq:RZ-Hamiltonian}]. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-splitting field. (b) Upper panel: a schematic illustration of Landau levels near the II-Dirac point of the Rashba model (with in-plane Zeeman field, and in the weak- hybridization limit). The solid line represents the Onsager-Lifshitz quantization of the trefoille orbit [cf.\ Eq.\ ({trefoillerule})], and the dashed line incorporates the first-order correction [cf.\ Eq.\ (\ref{RZfirstorderE})] due to tunneling. 
Lower panel: oscillation of $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})] for even integer $n$ with respect to magnetic field calculated by Eq. (\ref{RZfirstorderE}) (black dashed line) and numerical diagonalization (red solid line). $\delta E$ is extracted at the Landau level closest to 0.45eV and the II-Dirac point is located at 0.5eV. Parameters for lower panel of (b): $m{=}0.076m_0$, $\alpha{=} 1.5(10^{5}$cm/s$\cdot\hbar$), $Z_\parallel {=}$1meV. Out-of-plane Zeeman coupling is negligibly weak for the choice of $m$. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: an enlarged view of near II-Dirac point. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4{<}\epsilon_0{<}1.0$(eV).\cite{RZfignote}, Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$). Parameters for (d): $m=0.076m_0$, $l=100$\AA. \label{fig:RZ}}.
\end{figure}

@@

Away from the II-Dirac point, an electron undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of Landau-Zener tunneling/transmission and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_-}}{e^{i\tilde{\lambda}_+}}. \label{propinplanezeeman}}
The adiabatic phase $\tilde{\lambda}_{\pm}$ is the sum of a dynamical phase ${\pm} l^2 \delta S/2$ and  the single-band Berry phase $\phi^B_\pm$, for the outer ($+$) and inner ($-$) orbit respectively. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and decreasing (resp.\ increasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.
